Illuminating a Dark Lens : A Type Ia Supernova \\ Magnified by Galaxy Cluster Abell 2744

### 2015-05-01 authors.tex
incorporating minor comments on draft, post redshift change   * notably: updated references and minor rearrangement of section 5  and new text in 6.6
added AR
added and re-sorted authors, text modifications   * tweaks to discussion of free-form models and simulations

\newcommand{\Oxford}{Department of Physics, Oxford University, Keble Road, Oxford OX1 3RH, UK}

@@

@@

@@

\author{Marusa Bradac\altaffilmark{\affilref{UCDavis}}}
\affilreftxt{UCDavis}{\UCDavis}

\author{Jose~Maria~Diego\altaffilmark{\affilref{Cantabria}}}
\affilreftxt{Cantabria}{\Cantabria}

@@

@@

@@

@@

\author{Liliya Williams\altaffilmark{\affilref{Minnesota}}}
\affilreftxt{Minnesota}{\Minnesota}

>>>>

\newcommand{\Oxford}{Department of Physics, University of Oxford, Keble Road, Oxford OX1 3RH, UK}

@@

\newcommand{\INAF}{INAF, Osservatorio Astronomico di Bologna, via Ranzani 1, I-40127 Bologna, Italy}
\newcommand{\JPL}{Jet Propulsion Laboratory, California Institute of Technology, 4800 Oak Grove Drive, Pasadena, CA 91109, USA}
\newcommand{\INFN}{INFN, Sezione di Bologna, Viale Berti Pichat 6/2, I-40127 Bologna, Italy}

@@

\author{Maru\v{s}a Brada\v{c}\altaffilmark{\affilref{UCDavis}}}
\affilreftxt{UCDavis}{\UCDavis}

\author{Jose~M.~Diego\altaffilmark{\affilref{Cantabria}}}
\affilreftxt{Cantabria}{\Cantabria}

@@

@@

\author{Curtis McCully\altaffilmark{\affilref{LCOGT},\affilref{UCSB}}}
\affilreftxt{LCOGT}{\LCOGT}
\affilreftxt{UCSB}{\UCSB}

@@

\author{Massimo Meneghetti\altaffilmark{\affilref{INAF},\affilref{JPL},\affilref{INFN}}}
\affilreftxt{INAF}{\INAF}
\affilreftxt{JPL}{\JPL}
\affilreftxt{INFN}{\INFN}

@@

\author{Adam Riess\altaffilmark{\affilref{JHU},\affilref{STScI}}}
\affilreftxt{JHU}{\JHU}
\affilreftxt{STScI}{\STScI}

@@

\author{Liliya~L.~R.~Williams\altaffilmark{\affilref{Minnesota}}}
\affilreftxt{Minnesota}{\Minnesota}

### 2015-05-01 snTomas.tex
added and re-sorted authors, text modifications   * tweaks to discussion of free-form models and simulations

Galaxy clusters can be used as cosmic telescopes to magnify distant
background objects through gravitational lensing, which can
substantially increase the reach of deep imaging surveys.  The lensing
magnification enables the study of objects that would otherwise be
unobservable because they are either intrinsically
faint \citep[e.g.][]{Schenker:2012,Alavi:2014} or extremely
distant \citep[e.g.][]{Franx:1997,Ellis:2001,Hu:2002,Kneib:2004,Richard:2006,Richard:2008,Bouwens:2009a,Maizy:2010,Zheng:2012,Coe:2013,Bouwens:2014,Zitrin:2014b}.
Background galaxies are also {\it spatially} magnified, allowing for
studies of the internal structure of galaxies in the early universe
with resolutions of $\sim$100
pc \citep[e.g.]{Stark:2008,Jones:2010,Yuan:2011,Livermore:2015}.

@@

In general, free-form methods tend to explore the model parameter
space more fully than parametric methods.  This is reflected in the
size of the error-bars in Figure~\ref{fig:LensingTest}.  For example,
the Williams model (using GRALE) has the largest uncertainties, in
part because it accounts for mass sheet degeneracy by introducing a
free parameter in the reconstruction that represents an arbitrarily
scaled mass sheet, with normalization to be chosen by the genetic
algorithm.  More complex degeneracies are also accounted for in this
model through the use of basis
functions \citep{Liesenborgs:2006,Liesenborgs:2007,Mohammed:2014}

@@

A promising avenue for exploring the origins of systematic biases in
cluster lens models is through the use of simulated lensing data.  One can
start with very deep high-resolution multi-band imaging on an unlensed
field that has a fairly complete spectroscopic redshift catalog, such
as the Hubble Ultra Deep Field.  Then a simulated galaxy cluster is
placed in the field, and the background galaxies are distorted into
arcs and multiple images using a well-defined lensing prescription.
The artificially lensed images can then be distributed to lens
modeling teams who attempt to reconstruct the (known) mass profile of
the simulated cluster.  This exercise has recently been pursued with a
set of synthetic clusters similar to those observed in the HFF program
(Meneghetti et al. in prep).  Preliminary analysis of this simulation
comparison suggests that magnifications tend to be
overestimated for sources that lie outside the strong-lensing region
along the minor axis direction of an elongated cluster.  Our analysis
of SN \tomas\ here indicates that such simulation efforts will be an
important step for moving toward precision science with cluster-lensed
sources.

@@

Financial support for this work was provided to SAR by NASA through grants
HST-HF-51312 and HST-GO-13386 from STScI, which is operated by Associated Universities for
Research in Astronomy, Inc. (AURA), under NASA contract NAS 5-26555.

AM acknowledge the financial support of the Brazilian funding agency
FAPESP (Post-doc fellowship - process number 2014/11806-9)

Support for this research at Rutgers University was provided in part
by NSF CAREER award AST-0847157 to SWJ.  The Dark Cosmology Centre is
supported by the Danish National Research Foundation.

J.M.D acknowledges support of the consolider project CSD2010-00064 and
AYA2012-39475-C02-01 funded by the Ministerio de Economia y
Competitividad.

J.M. contributed to this research from the Jet Propulsion Laboratory,
California Institute of Technology, under a contract with NASA and
acknowledges support from NASA Grants HST-GO-13343.05-A and
HST-GO-13386.13-A. The research leading to these results has received
funding from the People Programme (Marie Curie Actions) of the European
Union's Seventh Framework Programme (FP7/2007-­2013) under REA grant
agreement number 627288.

A.Z. acknowledges financial support from NASA through grant
HST-HF-51334.01-A awarded by STScI and operated by AURA.

>>>>

Galaxy clusters can be used as cosmic telescopes to magnify distant
background objects through gravitational lensing, which can
substantially increase the reach of deep imaging surveys.  The lensing
magnification enables the study of objects that would otherwise be
unobservable because they are either intrinsically
faint \citep[e.g.][]{Schenker:2012,Alavi:2014} or extremely
distant \citep[e.g.][]{Franx:1997,Ellis:2001,Hu:2002,Kneib:2004,Richard:2006,Richard:2008,Bouwens:2009a,Maizy:2010,Zheng:2012,Coe:2013,Bouwens:2014,Zitrin:2014b}.
Background galaxies are also {\it spatially} magnified, allowing for
studies of the internal structure of galaxies in the early universe
with resolutions of $\sim$100
pc \citep[e.g.][]{Stark:2008,Jones:2010,Yuan:2011,Livermore:2015}.

@@

In general, free-form methods tend to explore the model parameter
space more fully than parametric methods.  This is reflected in the
size of the error-bars in Figure~\ref{fig:LensingTest}.  The Williams
model (using GRALE) has the largest uncertainties, \changeR{with a
68\% confidence range that encompasses the median magnification values
returned by all other models.  This is possibly the most complete
uncertainty estimate as it includes many systematic error sources
that are not incorporated into most other models. For example, this
implementation of GRALE} accounts for mass sheet degeneracy by
introducing a free parameter in the reconstruction that represents an
arbitrarily scaled mass sheet, with normalization to be chosen by the
genetic algorithm.  More complex degeneracies are also accounted for
in this model through the use of basis
functions \citep{Liesenborgs:2006,Liesenborgs:2007,Mohammed:2014}.

@@

A promising avenue for exploring the origins of systematic biases in
cluster lens models is through the use of simulated lensing data.  One
can start with very deep high-resolution multi-band imaging on an
unlensed field that has a fairly complete spectroscopic redshift
catalog, such as the Hubble Ultra Deep Field.  Then a simulated galaxy
cluster is placed in the field, and the background galaxies are
distorted into arcs and multiple images using a well-defined lensing
prescription.  The artificially lensed images can then be distributed
to lens modeling teams who attempt to reconstruct the (known) mass
profile of the simulated cluster.  This exercise has recently been
pursued with a set of synthetic clusters similar to those observed in
the HFF program (Meneghetti et al. in prep).  Preliminary analysis of
this simulation comparison suggests that magnifications tend to be
overestimated for sources that lie outside the strong-lensing region
along the minor axis direction of \changeR{simulated clusters that are
very elongated, similar to Abell 2744. This finding would be
consistent with a recent lens model comparison on 25 real clusters
from the CLASH program \citep{Zitrin:2015}.} Our analysis of
SN \tomas\ here indicates that such simulation efforts \changeR{and
head-to-head lens model comparisons} will be an important step for
moving toward precision science with cluster-lensed sources.

@@

Financial support for this work was provided to S.A.R. by NASA through
grants HST-HF-51312 and HST-GO-13386 from STScI, which is operated by
Associated Universities for Research in Astronomy, Inc. (AURA), under
NASA contract NAS 5-26555. A.M. acknowledge the financial support of
the Brazilian funding agency FAPESP (Post-doc fellowship - process
number 2014/11806-9). Support for this research at Rutgers University
was provided in part by NSF CAREER award AST-0847157 to SWJ.  The Dark
Cosmology Centre is supported by the Danish National Research
Foundation. J.M.D acknowledges support of the consolider project
CSD2010-00064 and AYA2012-39475-C02-01 funded by the Ministerio de
Economia y Competitividad. J.M. contributed to this research from the
Jet Propulsion Laboratory, California Institute of Technology, under a
contract with NASA and acknowledges support from NASA Grants
HST-GO-13343.05-A and HST-GO-13386.13-A. The research leading to these
results has received funding from the People Programme (Marie Curie
Actions) of the European Union's Seventh Framework Programme
(FP7/2007-­2013) under REA grant agreement number
627288. A.Z. acknowledges financial support from NASA through grant
HST-HF-51334.01-A awarded by STScI and operated by AURA. LLRW
acknowledges the support of the Minnesota Supercomputing Institute.
